Assessing the Population-Level Conservation Effects of Marine Protected Areas

### 2020-11-30 documents/referee-replies.Rmd

XX methods section XX

*Line 146: all this paragraph should be in the introduction*

We respectfully disagree with this suggestion. Following referee #1's helpful suggestions, we have clarified the broader challenges with interpreting response ratios as population-level effects in the introduction on line XX, which we agree were poorly articulated. However, we feel that this section exploring the specific response ratios in the Channel Islands is a result of our simulation modeling, and so belongs in the results section. Moving this paragraph to the introduction would require explaining our simulation modeling in the introduction as well, which we feel is not ideal.

XX Owen, thoughts on tone here?XX
XX Looks great to me!--O XX

@@

*Line 50: please precise what do you mean by “alteration in fishing effort”*

Thank you for pointing this out. We have removed that line following the Referee #1's suggestion to ensure to continuity of paragraphs, as the displacement of fishing effort is not directly a population level effect, but rather a cause of population effect.

*Line 51-52: “Why does this large body of empirical work not serve as sufficient evidence for the population-level effects of MPAs?” In my opinion, you did not define clearly why all the literature you cited before does not provide sufficient evidence for the population-level effects of MPAs. You just referred to publication but you did not explain why the results are debatable.*

Per the referees helpful suggestion, we have moved materials related to this from the results to the introduction (line XX). The problem is that response ratio / spatial BACI style designs will still be biased if the treatment unit affects the control unit, a violation of the stable unit treatment value assumption (SUTVA). For example, suppose we employ a spatial BACI/BACIPS design on a system with two patches of equal size, one of which is put in a perfectly enforced MPA, the other left open to fishing. Suppose then that all the fishing effort that used to operate inside the MPA moves and now operates outside the MPA, doubling the amount of fishing pressure outside. A BACI design that assumes that the control site is a measure of what would have happened in the absence of an MPA will conclude that the MPAs had a large positive effect, since biomass outside the MPA will decline post-MPA, while biomass inside the MPA will increase. Conversely, consider that same MPA but without effort displacement but with very strong larval and adult spillover, which results in the control site benefit greatly from the MPA. A BACI style design would conclude that the MPA had little to no effect in this case, since biomass levels would be similar inside and outside the MPA.

XX Owen check if you think this is clearer now. basically most fo the intro XX
XX This explanation is good and I have made minor edits in the intro XX

@@

The posterior probability distribution is what allows us to interpret the say 95% credible interval of a parameter as "there is a 95% probability that conditional on the data, model, and priors the effect of Z was between X and Y". In contrast a frequentist 95% confidence interval tells us that there is a 95% change that if were to resample the data infinite times and refit the model, there is a 95% chance that the true parameter value would fall within that range.

So, in this case, we are reporting the posterior probability distribution of the MPA effect, conditional on the data (and the other covariates etc, priors, and model structure).

Given the space constraints of this paper, we do not feel that we are able to include a broader discussion of Bayesian inference, but have included a citation to a standard textbook on the topic in our introduction of our methods on line XX.

XX Owen, too snarky? XX
XX Thisf response is fine, and I will check the brief description in the main text as well XX

### 2020-11-30 documents/ovando-et-al-pop-effects-of-mpas.Rmd
pre-chop-back-commit

No-take Marine Protected Areas (MPAs), spatial regions of the ocean in which fishing is prohibited, have a long history in the management of marine resources. Modern MPAs were first established as marine analogs to the terrestrial protection of iconic landscapes [@iucn1976]. Over time our goals and expectations for MPAs have evolved and become more ambitious. Recent international efforts to expand MPAs, such as the Biodiversity Beyond National Jurisdiction (BBNJ) Agreement and the Convention on Biological Diversity’s Post-2020 Global Biodiversity Framework, are based in part on the assumption that well-designed MPAs will not only provide conservation benefits inside their borders, but will have broader conservation effects on unprotected areas surrounding the MPAs[@gaines2010].

The empirical MPA literature has focused on assessing the ability of MPAs to provide conservation gains within their borders [@lester2009; @edgar2014; @gerber2005]. However, as conservation benefits accrue inside MPAs, MPAs also affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort. Therefore, MPAs contribute to both local and population-level effects (i.e., inside-MPA effects and broader regional effects). Theory states that numerous factors influence how MPAs might affect fish populations. These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], the strength, timing, and location of density dependence [@burgess2014], the degree of enforcement [@edgar2014], the level of fishing pressure, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006a;@walters2000a;@white2011,;@white2010c;@ovando2016;@moffitt2013]. The location and spacing of the MPAs within a network can also influence their population-level impacts through synergistic effects, for example acting as stepping stones for larvae [@gaines2010; @costello2010a].

<!-- What empirical evidence is there for these population-level effects of MPAs?  -->
These theoretical results are generally based on modeling of closed populations, where some fraction of the connected population is protected inside of MPAs. In contrast to this population paradigm used in MPA simulations, MPAs are often evaluated empirically at local scales using "response ratios", commonly measured as the ratio of biomass densities (weight of organisms per unit area) of species inside relative to selected control sites outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015]. These studies found clear evidence that well enforced and sufficiently sized MPAs are associated with high response ratios. Several studies have also documented empirical evidence for the existence of adult or larval fish spillover affecting fish abundance [@halpern2009;@kay2012; @stobart2009; @mcclanahan2000; @russ1996; @thompson2017;@baetscher2019;@pelc2009]. Where response ratios are available before and after MPA implementation, before-after-control-impact (BACI) style studies have shown similarly clear and positive results [@thiault2019a].

While these studies have clearly demonstrated the ability of MPAs to create differences between local fished and unfished areas, this body of empirical work does not serve as sufficient evidence for the broader, population-level effects of MPAs. Spillover-based studies provide evidence that MPAs may affect areas outside their boundaries, but these studies often lack a counterfactual, and so cannot determine the magnitude of population-level effects. Response ratios have shortcomings as measures of population-level MPA effects as well. Control sites used in calculating response ratios are often selected based on habitat characteristics [@ferraro2018]. However, export of adults or larvae from the MPA to these "control" sites affects their status as controls, as does displacement of fishing effort from MPAs to control sites. These factors can act to bias estimators that require the assumption that the control sites are unaffected by the treatment (a required assumption of response ratios and BACI designs [@ferraro2018;@moffitt2013]). In theory, control sites far enough away to negate both spillover of fish or larvae, and concentration of the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs is challenging. As variations of response ratios are our primary source of evidence for the conservation effects of MPAs, this means that our collective empirical understanding of the population-level impacts of MPAs is surprisingly limited.

@@

We present here an paired theoretical and empirical assessment demonstrating the challenges in assessing the population-level impacts of MPAs. In 2003, a network of MPAs was established in the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) covering approximately 20% of the Islands' waters [@kirlin2013]. The network has been used as a model case study in protected area design around the world (@botsford2014 and references therein). We use data from the first 14 years of protection in a difference-in-difference model [see @angrist2009 and references therein] to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of fin-fish species. In contrast to clear differences in biomass densities observed inside and outside of well-protected MPAs both globally [@lester2009] and in the Channel Islands [@caselle2015] we are unable to detect a clear population effect from the Channel Islands MPAs. We build off of existing MPA theory to consider why this might be, and provide guidance for scientists and managers as to when and how we might expect to estimate the population-level conservation effects of MPAs.

@@

These simulation results provide a library of plausible MPA effects for a range of biological and economic assumptions. One set of simulations is specifically designed to reflect the dynamics of the Channel Islands. We only include species of the same genus as those targeted by fishing in the Channel Islands, restrict fishing pressure such that the simulated populations are moderately to lightly exploited (given the best available knowledge of stock status for the targeted species in the Channel Islands), and cap the MPA size at 20% of the population's range [@rassweiler2012]. For each of these Channel Island simulations, we calculated biomass density response ratios.

@@

For readers more familiar with BACI methods, the difference-in-difference (DiD) regression amounts to estimating the pre-post MPA difference in the biomass densities of targeted species, minus the same difference for non-targeted species in the Channel Islands.

@@

This difference-in-difference regression provides our estimate of the population-level conservation effect of the Channel Islands MPAs. We use a Bayesian regression since it allows us to interpret our estimated effects probabilistically [@mcelreath2020], e.g. stating the probability that conditional on the model, priors,and data the Channel Island MPAs increased or decreased biomass of targeted species. Being a Bayesian regression, our difference-in-difference model produces posterior probability distributions of our coefficients, from which we can construct Bayesian credible intervals. Since the effects of MPAs will evolve over time, we use this model to estimate a series of population-level effects of the Channel Islands MPAs in three-year intervals from 2003 to 2017.

To compare response ratios from the Channel Islands simulations to our empirical results, we organized simulated and empirical response ratios into bins. Because each simulation included measures of both response ratios and population-level effects, this provided a library of simulations (and their associated attributes) that could have produced the types of empirical response ratios measured in the Channel Islands.

@@

These response ratios cannot, however, be used as a definitive indicator of population effects of the MPAs. In the case of the Channel Islands MPAs, control sites are often located within a few kilometers of an MPA, suggesting that they are susceptible to both local biological spillover and by concentration of fishing effort excluded from the MPAs. To illustrate this problem, we paired our simulated response ratios to the estimated posterior probability distributions of the empirical response ratios by matching the patterns in the empirical response ratios. The response ratio trends we observe in the data could plausibly be produced by a wide range of population-level MPA effects (Fig.\@ref(fig:rr-plot)). Response ratios well over one were associated with "true" population-level MPA effects generally less than 25%, and, importantly, many simulations produced large response ratios but population-level MPA effects close to 0%. This can occur if for example fishing pressure is only moderate, adult movement is low, larval dispersal is high, and displaced fishing effort concentrates around the border of the MPAs.

@@

Our empirical difference-in-difference regression provides a different perspective on population-wide MPA effects. Over the first three years of implementation (2003-2006), the effects of the MPAs are unclear, with support for a small negative effect to a substantially positive effect, with much higher probability of a small positive effect (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`,
90\% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`
- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`). Over the next six years the model estimates greater probabilities of an increasingly positive MPA effect, peaking in 2009-2011 with a median estimate of MPA effect of a `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)` increase in mean total biomass density of targeted species (90% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)`
- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)`) (Fig.\@ref(fig:mpa-effect)). These empirical estimates are in line with the outcomes that our simulation model suggests are plausible. However, in the subsequent years the trend reverses itself, and in 2015-2017 we once again see no clear effect of the MPAs (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`,
90\% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`
- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`.) Fig.\@ref(fig:mpa-effect).

Given natural variability in ecosystems and uncertainty in measurements, MPA effects may be difficult to detect with precision. Our library of simulation results allows us to explore how MPA effects, and our ability to measure them, might vary with observation error and natural fish recruitment variation (including the possibility of recruitment regimes). Using a Bayesian difference-in-difference regression on these simulated data, we estimated the percent error between the posterior probability distribution of the estimated MPA effect from the regression and the true simulated MPA effect.

The model struggled severely when MPA effect sizes were less than 25% and the model was faced with both observation and process errors (Fig.\@ref(fig:val-plot)). Even models fit to data generated from large effect sizes commonly mis-estimated the true MPA effect by 50% or more.  Obtaining a mean absolute percent error (MAPE) of 25% or less across our simulated datasets required a "true" population-level MPA effect of at least 30%. In the context of the Channel Islands, this finding suggests that we should not be surprised at our difficulty in precisely estimating the population-level effect of the MPAs.
<!-- This is merely an illustrative exercise, omitting critical factors such as detection probability and sampling strategy. However, since nearly any omission which one can think of would make an MPA effect harder to detect, not easier, these results serve as a useful floor for the likely difficulty in estimating MPA effects. -->

@@

Containing a carefully designed, well-enforced, and well-studied MPA network, the Channel Islands seems to be an ideal location to study the population-level effects of protected areas. Persistently high response ratios suggest that despite overall decreases in targeted biomass densities inside and outside MPAs relative to pre-MPA measurements (Fig.\@ref(fig:raw-density)), the MPAs may still be providing protection within their borders. But, these response ratios are not necessarily an indicator of broader population-level effects. The difference-in-difference strategy presents an alternative identification strategy. It is not without its own strict caveats, but provides some potential improvements over response-ratios as a means of estimating population-level effects. While we estimate an uncertain but overall positive effect of the MPA network in its first few years, we are unable to detect a robust positive signal from 2012-2017. After 14 years of MPA protection we are left without a clear picture of the effect of the Channel Island MPA network on biomass densities of targeted fin-fish species from either response ratios or the difference-in-difference model.

Uncertain fishing dynamics may be one factor contributing to a lack of strong MPA network effects. Much of the theoretical literature on MPAs assumes that larger reserves produce larger conservation gains [@white2011 and references therein]. However, these models generally simulate fleet dynamics through fishing mortality rates, the proportion of total mortality experienced by a population attributable to fishing pressure [e.g. @halpern2004]. The implication of this is that total catches will scale with the size of the available population.

@@

While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (SI Fig.S1) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely. However, some of the most common targeted species in the Channel Islands, such as blue and copper rockfish, have seen dramatic increases in reported catches in the Santa Barbara region since the year 2010 (although other such as California sheephead have seen declines relative to pre-MPA catch level). While we include catch histories in our regression, to the extent that these increases in catches are due to external market forces, the disappearance of clear positive population-level MPA effects estimated by our model could be due to exogenous increases in fishing pressure overwhelming any population-level increases caused by the MPA network.

@@

MPAs are an important part of the marine resource management toolbox. As the number and size of global MPA networks increase, we must set appropriate expectations for their outcomes on both local and regional scales. Our results for the relatively large, well-enforced, and rigorously studied (though also relatively lightly fished) Channel Islands Marine Protected Area network make clear that in many instances empirically detecting a clear effect of MPAs on total biomass densities of targeted fin-fish at the population level may not be possible. In the absence of clear empirical findings, simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can be tuned to focus on the species groups that have the highest chance of a detectable effect size over the early years of the reserve. Expanding data collection to include robust monitoring of spatio-temporal fleet dynamics may help assess the validity of control sites used in response ratios, support the direct inclusion of these fleet dynamics into statistical models, and allow managers to take into account potential negative interactions between MPAs and fleet dynamics such as those that may occur under constant-catch dynamics. Whenever possible monitoring programs should be implemented prior to MPA implementation to provide a pre-treatment benchmark.

>>>>

No-take Marine Protected Areas (MPAs), spatial regions of the ocean in which fishing is prohibited, have a long history in the management of marine resources. Modern MPAs were first established as marine analogs to the terrestrial protection of iconic landscapes [@iucn1976]. Over time our goals and expectations for MPAs have evolved and become more ambitious. Recent international efforts to expand MPAs, such as the Biodiversity Beyond National Jurisdiction (BBNJ) Agreement and the Convention on Biological Diversity’s Post-2020 Global Biodiversity Framework, are based in part on the assumption that well-designed MPAs will not only provide conservation benefits inside their borders, but also have broader conservation effects on unprotected areas surrounding the MPAs [@gaines2010].

The empirical MPA literature has focused on assessing the ability of MPAs to provide conservation gains within their borders [@lester2009; @edgar2014; @gerber2005]. However, as conservation benefits accrue inside MPAs, MPAs also affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort. Therefore, MPAs contribute to both local and population-level effects (i.e., inside-MPA effects and broader regional effects). Numerous factors influence how MPAs affect fish populations. These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], the strength, timing, and location of density dependence [@burgess2014], the design of the network [@gaines2010; @costello2010a],the degree of enforcement [@edgar2014], the level of fishing pressure, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006a;@walters2000a;@white2011;@white2010c;@ovando2016;@moffitt2013].

<!-- What empirical evidence is there for these population-level effects of MPAs?  -->
These theoretical results are generally based on modeling of closed populations, where some fraction of a population is protected inside of MPAs. In contrast to this population paradigm used in MPA simulations, MPAs are often evaluated empirically at local scales using spatial "response ratios", commonly measured as the ratio of biomass densities (weight of organisms per unit area) of species inside relative to selected control sites outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015]. These studies found clear evidence that well enforced and sufficiently sized MPAs are associated with high response ratios. Several studies have also documented empirical evidence for the existence of adult or larval fish spillover affecting fish abundance [@halpern2009;@kay2012; @stobart2009; @mcclanahan2000; @russ1996; @thompson2017;@baetscher2019;@pelc2009]. Where response ratios are available before and after MPA implementation, spatial before-after-control-impact (BACI) style studies have shown similarly clear and positive results [@thiault2019a].

While these studies have clearly demonstrated the ability of MPAs to create differences between local fished and unfished areas, this body of empirical work does not necessarily serve as sufficient evidence for the population-level effects of MPAs. This is because these studies rely on assuming that selected control sites serve as a measure of what would have happened in the absence of MPAs. Control sites used in calculating response ratios are often selected based on habitat characteristics to justify their use as counterfactuals [@ferraro2018]. However, the very spillover effects MPAs are increasingly hoped to produce can negate the ability of spatial response ratio / BACI designs to accurately estimate the effects of MPAs. Export of adults or larvae from MPAs to "control" sites can affect their status as controls, as does displacement of fishing effort from MPAs to control sites. These factors can bias estimators of population-level effects that require the assumption that the control sites are unaffected by the treatment (a required assumption of response ratios and BACI designs [@ferraro2018;@moffitt2013]). Control sites sufficiently far from MPAs to negate both spillover of fish or larvae and concentration of the fishing fleet could be selected, but finding suitably distant sites that are also appropriate proxies for the ecological and economic context of the MPAs is challenging. As variations of spatial response ratios and BACI studies are our primary source of evidence for the conservation effects of MPAs, this means that our empirical understanding of the population-level impacts of MPAs is surprisingly limited.

@@

We present here an paired theoretical and empirical assessment demonstrating the challenges in assessing the population-level impacts of MPAs. In 2003, a network of MPAs was established in the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) covering approximately 20% of the Islands' waters [@kirlin2013]. The network has been used as a model case study in protected area design around the world (@botsford2014 and references therein). We use data from the first 14 years of protection in a difference-in-difference model [see @angrist2009 and references therein] to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of fin-fish species. Rather than relying on spatial controls as is typically done in MPA studies, we use groups of species targeted and not-targeted by fishing pressure as our treatment and control groups. In contrast to clear differences in biomass densities observed inside and outside of well-protected MPAs both globally [@lester2009] and in the Channel Islands [@caselle2015] we are unable to detect a clear population effect from the Channel Islands MPAs. We build off of existing MPA theory to consider why this might be, and provide guidance for scientists and managers as to when and how we might expect to estimate the population-level conservation effects of MPAs.

@@

These simulation results provide a library of plausible MPA effects for a range of biological and economic assumptions. One set of simulations is specifically designed to reflect the dynamics of the Channel Islands. We only include species of the same genus as those targeted by fishing in the Channel Islands, restrict fishing pressure such that the simulated populations are moderately to lightly exploited (given the best available knowledge of stock status for the targeted species in the Channel Islands), and cap the MPA size at 20% of the population's range [@rassweiler2012]. For these Channel Island-esque simulations, we calculated 1) the population-level response for each species (the 'true' population-wide difference in biomass between the simulations with and without the MPAs) and 2) the measured response ratio of biomass densities inside and outside the simulated MPAs. To compare response ratios from the Channel Islands simulations to our empirical results, we organized simulated and empirical response ratios into bins. Because each simulation included measures of both response ratios and population-level effects, this provided a library of simulations (and their associated attributes) that could have produced the types of empirical response ratios measured in the Channel Islands.

@@

The difference-in-difference (DiD) regression amounts to estimating the pre-post MPA difference in the biomass densities of targeted species, minus the same difference for non-targeted species in the Channel Islands.

@@

This difference-in-difference regression provides our estimate of the population-level effect of the Channel Islands MPAs. We use a Bayesian regression since it allows us to interpret our estimated effects probabilistically [@mcelreath2020], e.g. stating the probability that conditional on the model, priors,and data the Channel Island MPAs increased or decreased biomass of targeted species. Being a Bayesian regression, our difference-in-difference model produces posterior probability distributions of our coefficients, from which we can construct Bayesian credible intervals. Since the effects of MPAs will evolve over time, we use this model to estimate a series of population-level effects of the Channel Islands MPAs in three-year intervals from 2003 to 2017.

@@

These response ratios cannot, however, be used as a definitive indicator of population effects of the MPAs. In the case of the Channel Islands MPAs, control sites are often located within a few kilometers of an MPA, suggesting that they are susceptible to both local biological spillover and by concentration of fishing effort excluded from the MPAs. Our simulation results show that the response ratio trends we observe in the data could plausibly be produced by a wide range of population-level MPA effects (Fig.\@ref(fig:rr-plot)). Response ratios well over one were associated with "true" population-level MPA effects generally less than 25%, and, importantly, many simulations produced large response ratios but population-level MPA effects close to 0%. This can occur if for example fishing pressure is only moderate, adult movement is low, larval dispersal is high, and displaced fishing effort concentrates around the border of the MPAs.

@@

Our difference-in-difference regression provides a different perspective on population-wide MPA effects. Over the first three years of implementation (2003-2006), the effects of the MPAs are unclear, with support for a small negative effect to a substantially positive effect, with much higher probability of a small positive effect (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`,
90\% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`
- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`). Over the next six years the model estimates greater probabilities of an increasingly positive MPA effect, peaking in 2009-2011 with a median estimate of MPA effect of a `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)` increase in mean total biomass density of targeted species (90% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)`
- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)`) (Fig.\@ref(fig:mpa-effect)). These empirical estimates are in line with the outcomes that our simulation model suggests are plausible. However, in the subsequent years the trend reverses itself, and in 2015-2017 we once again see no clear effect of the MPAs (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`,
90\% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`
- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`.) Fig.\@ref(fig:mpa-effect).

Given natural variability in ecosystems and uncertainty in measurements, MPA effects may be difficult to detect with precision. Our library of simulation results allows us to explore how MPA effects, and our ability to measure them, might vary with observation error and natural fish recruitment variation (including the possibility of recruitment regimes). Using a Bayesian difference-in-difference regression on these simulated data, we estimated the percent error between the posterior probability distribution of the estimated MPA effect from the regression and the true simulated MPA effect. The model struggled severely when MPA effect sizes were less than 25% and the model was faced with both observation and process errors (Fig.\@ref(fig:val-plot)). Even models fit to data generated from large effect sizes commonly mis-estimated the true MPA effect by 50% or more. Obtaining a mean absolute percent error (MAPE) of 25% or less across our simulated datasets required a "true" population-level MPA effect of at least 30%. In the context of the Channel Islands, this finding suggests that we should not be surprised at our difficulty in precisely estimating the population-level effect of the MPAs.
<!-- This is merely an illustrative exercise, omitting critical factors such as detection probability and sampling strategy. However, since nearly any omission which one can think of would make an MPA effect harder to detect, not easier, these results serve as a useful floor for the likely difficulty in estimating MPA effects. -->

@@

Containing a carefully designed, well-enforced, and well-studied MPA network, the Channel Islands seems to be an ideal location to study the population-level effects of protected areas. Persistently high response ratios suggest that despite overall decreases in targeted biomass densities inside and outside MPAs relative to pre-MPA measurements (Fig.\@ref(fig:raw-density)), the MPAs may still be providing protection within their borders. But, these response ratios are not necessarily an indicator of broader population-level effects. The targeted vs. non-targeted difference-in-difference strategy presents an alternative means of estimating the population-level effects of MPAs. While having its own strict caveats, this strategy provides some potential improvements over spatial response-ratios as a means of estimating population-level effects. While we estimate an uncertain but overall positive effect of the MPA network in its first few years, we are unable to detect a robust positive signal from 2012-2017. After 14 years of MPA protection we are left without a clear picture of the population-level effect of the Channel Island MPA network on biomass densities of targeted fin-fish species.

Fishing dynamics may be one factor contributing to a lack of strong MPA network effects. Much of the theoretical literature on MPAs assumes that larger reserves produce larger conservation gains [@white2011 and references therein]. However, these models generally simulate fleet dynamics through fishing mortality rates; the proportion of total mortality experienced by a population attributable to fishing pressure [e.g. @halpern2004]. The implication of this is that total catches will scale with the size of the available population.

@@

While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (SI Fig.S1 XX ) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely. However, some of the most common targeted species in the Channel Islands, such as blue and copper rockfish, have seen dramatic increases in reported catches in the Santa Barbara region since the year 2010 (although other such as California sheephead have seen declines relative to pre-MPA catch level). While we include catch histories in our regression, to the extent that these increases in catches are due to external market forces, the disappearance of clear positive population-level MPA effects estimated by our model could be due to exogenous increases in fishing pressure overwhelming any population-level increases caused by the MPA network.

@@

<!-- MPAs are an important part of the marine resource management toolbox. As the number and size of global MPA networks increase, we must set appropriate expectations for their outcomes on both local and regional scales. Our results for the relatively large, well-enforced, and rigorously studied (though also relatively lightly fished) Channel Islands Marine Protected Area network make clear that in many instances empirically detecting a clear effect of MPAs on total biomass densities of targeted fin-fish at the population level may not be possible. In the absence of clear empirical findings, simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can be tuned to focus on the species groups that have the highest chance of a detectable effect size over the early years of the reserve. Expanding data collection to include robust monitoring of spatio-temporal fleet dynamics may help assess the validity of control sites used in response ratios, support the direct inclusion of these fleet dynamics into statistical models, and allow managers to take into account potential negative interactions between MPAs and fleet dynamics such as those that may occur under constant-catch dynamics. Whenever possible monitoring programs should be implemented prior to MPA implementation to provide a pre-treatment benchmark. -->
